Wikipedia WALS Feature Analysis

### 2019-07-29 surprisal/WALS_features.Rmd
geographic distance between languages

@@

imputed_df <- wals %>% 
  select(-`139A`, -`141A`, -`140A`) %>% # these features have no values
  select(language, `121A`:`120A`) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>% 
  MIMCA(ncp = 2, threshold = 1e-02)

imputed_df110 <- as.data.frame(imputed_df[[1]][[99]]) 
```

@@

wf_sub <- all_imputed %>% 
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))

>>>>

```{r geographic area comparison}
languages <- read_csv(here("Data/languages.csv")) %>%
  select(Latitude, Longitude, ISO639P3code)

wiki <- read_csv(here("Data/relative_ngrams.csv"))

labs <- read_csv(here("Data/wiki_wals.csv")) %>% 
  group_by(language) %>% 
  slice(1) %>% 
  ungroup() %>% 
  select(language, iso)

latlong <- labs %>% 
  left_join(imputed_df) %>% 
  left_join(languages, by = c("iso" = "ISO639P3code")) %>%
  group_by(language) %>%
  slice(1) %>% 
  ungroup() %>%
  select(language, Latitude, Longitude) %>%
  filter(complete.cases(.))

sub_langs <- latlong %>% 
  pull(language)

unigram_cosines <- wiki %>%
  filter(language %in% sub_langs, gram == "Unigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, unigram_cosine, -language1)

wf_sub <- imputed_df %>% 
  filter(language %in% sub_langs) %>%
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))

compute_geo_sim <- function(language1, language2) {

  latlong %>%
    filter(language == language1) %>%
    bind_cols(latlong %>% filter(language == language2)) %>%
    mutate(geo = sqrt((Latitude1 - Latitude)^2 + (Latitude1 - Latitude)^2)) %>%
    ungroup() %>%
    rename(language2 = language1, language1 = language) %>% 
    select(language1, language2, geo)
}

geo_space <- map_dfr(1:nrow(wf_pairwise), ~compute_geo_sim(wf_pairwise[.x,"language1"],
                                          wf_pairwise[.x,"language2"]))

@@

imputed_df <- wals %>% 
  select(-`139A`, -`141A`, -`140A`) %>% # these features have no values
  select(language, `121A`:`120A`) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>% 
  MIMCA(ncp = 2, threshold = 1e-02)
```

@@

wf_sub <- imputed_df %>% 
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))
